### 2009-08-28 latex/paper1/contents.tex
paper

Unfortunately, in the case of legacy systems, such knowledge isn't available.
Depending on its size, it might take months to understand the system so well as
to find a good modularization.

@@

@@

\section{Software Networks and Software Clustering}

@@

\section{Models}

@@

@@

We propose an extension to BCR model...

@@

First of all, the model generates a network that is a copy of the module
dependency network, each vertex belonging to its respective module. Then the
algorithm runs iteratively, and on each iteration it picks a event that evolves
the network. Each event has an associated probability.

TODO: present in pseudo-language?

* With probability p, a new vertex is added to the network, together with an
edge from the new vertex to an existing vertex, chosen preferentially according
to .... Considering din = 0, this means that a vertex with indegree = 8 is
twice more likely to receive the edge than a vertex with indegree = 4. The
parameter din can be used to alleviate the handicap. (If din = 4, the first
vertex will only be 3/2 more likely to receive the edge). The new vertex is
put on the same cluster as the old vertex.
* With probability q, a new vertex is added to the network, together with an
edge from an existing vertex, chosen preferentially accordingly to dout +
outdegree, to the new vertex. The new vertex is put on the same cluster as the
old vertex
* With probability r, a new edge is added between two vertices, v and w. v is
chosen according to .... After that, with prob X, w is chosen from one of the
modules connected to v's module, else it's chosen from v's own module. In any
case, the exact choice is made according to ...

@@

@@

@@

We want to show investigate if the models can produce networks that resemble
software networks. We know that they share with software networks the scale-free
property. This is not enough, since many real networks share this property. So
we looked for a method to differentiate between software networks and other
networks.

What we are looking for, after all, is an oracle that accepts software networks
and rejects non-software networks. If the oracle has these two properties, we
can be confident that it'll accept only synthetic networks that resemble
software networks.

In a recent work, Milo et al. proposed the study of triads in order to
characterize different classes of networks. We thus follow their work here.

Triads are...

Figure 1a show triads for a software system... Figure 2a show triads for network
from domain X.

\subsection{Software Network Extraction}

We've collected 65 systems written in Java and X networks from many domains,
such as sociology, biology, technology and linguistics.
Table 1 ...

\subsection{Oracle Training and Evaluation}

We then used Pearson's correlation coefficient as a similarity measure between
two networks. For each software network, We then computed the average
correlation coefficient (ACC) to the other software networks. We've observed
that among software networks the ACC is X +- Y.

We then computed, for each non-software network, the ACC to the 65 software
networks. By the 3-sigma rule, we use X as the threshold for realistic software
networks: networks whose ACC is below this threshold are rejected.

The oracle has X precision and X recall...

@@

We chose five different module dependency networks, which where extracted from
actual dependencies between modules of five different software systems of our
sample, ranging from 2 to 32 modules. The module dependency networks are shown
on Figure XXX.

@@

%02-gef
%04-ibatis
%08-megamek
%16-findbugs
%32-zk

In order to asses the realism of the models we should test them with

In order to assess the realism of the models, we generated a large number of
networks using the models with many combinations of parameters.

We expect that,
if a model is capable of producing a realistic network, then

We generated networks with many combinations of parameters...

BCR: 5 different architectures, 
p, q, r in (0.0, 0.2, 0.4, 0.6, 0.8, 1.0), with p+q+r = 1 and p + q > 0.
...
Total: X networks

LR: ...

CGW: ...

ER: ...

\section{Experimental Results}

Table: Model | Number of Networks | Realistic Networks | \%

@@

As expected, ER produces non-realistic software networks.

@@

Pick realistic networks from a model. Are they similar to each other? (see
standard deviation) Are they similar to networks generated by the other models?

\section{Threads to validity}

@@

We generated only one network for each set of parameters.

@@

# Koschke says that experts decompositions vary by no more than 80? percent.

@@

The use of models lead to a more controlled way of experimented, and allows us
to play with a large data set and a wide variety of characteristics. Although
we have shown that the models studied can produce realistic networks according
to one criterium, we can't just discard the experimentation with real software
systems, as they are the subjects in which algorithms will be used in the
real world. In particular, it's an open question whether the models can 
reproduce any software system or if they just generate a subset of systems.
We believe that there are software systems that cannot be simulated by none
of these models.

Although common in distributed computing research, simulation is underexplored
in software engineering. This work opens a door to the usage of simulation in
the field of reverse engineering of software in order to evaluate RE 
algorithms.

Future work: to apply algorithms to the networks and compare the results with
results found in the literature.

@@

Name | Source

@@

Description | Source

>>>>

Unfortunately, in the case of legacy systems, such knowledge isn't available.
Depending on its size, it might take months to understand the system so well as
to find a good modularization. % CITE Parnas' modules

@@

% TODO: diagram showing the evaluation of a clustering algorithm

@@

\section{Software Modularization Recovery}

Aka software architecture recovery, software architecture reconstruction,
software clustering...

It's an task of reverse engineering, as defined by Tonella...

%\section{Software Networks and Software Clustering}

@@

(in fact we can't argue the degree distribution is perfectly fit by a power law.
Anyway, the distribution is much more assymetric than normal ou Poisson's)

software indegree distribution (power law), outdegree distribution (not so power
law)

\section{Network Models}

@@

%\subsection{Configuration model}
% take the most typical software system and steal its degree sequence

@@

We propose an extension to BCR model... which was proposed to model the links
between web pages.

@@

First of all, the model generates a network that is a copy of the module
dependency network, each vertex belonging to its respective module. Then the
algorithm runs iteratively, and on each iteration it executes an operations that
modifies the network. It can add a new vertex together with an edge connecting it
to an old vertex, or it can add an edge between two existing vertices. The
choice of vertices, though, is not fully random. The probability that a
particular vertex v is choosen, P(v), is proportional to a function that involves the
in-degree or the out-degree of the node. We say that a vertex is chosen within a
set V according to a function f if

P(choose v) = f(v) / sum(v in V, f(v))

The denominator is a normalizing factor that assures that the probabilities sum
to 1.

picks a event that evolves
the network. Each event has an associated probability.

* Add a node with ongoing edge. With probability p, a new vertex is added to the network, together with an
edge from the new vertex to an existing vertex, chosen preferentially according
to .... Considering din = 0, this means that a vertex with indegree = 8 is
twice more likely to receive the edge than a vertex with indegree = 4. The
parameter din can be used to alleviate the handicap. (If din = 4, the first
vertex will only be 3/2 more likely to receive the edge). The new vertex is
put on the same cluster as the old vertex.

* Add a node with ingoing edge. With probability q, a new vertex is added to the network, together with an
edge from an existing vertex, chosen preferentially accordingly to dout +
outdegree, to the new vertex. The new vertex is put on the same cluster as the
old vertex. This case is similar to the previous case.

* Add an edge. With probability r, a new edge is added between two vertices, v and w. v is
chosen according to .... After that, with prob X, w is chosen from one of the
modules connected to v's module, else it's chosen from v's own module. In any
case, the exact choice is made according to ...

@@

The CGW was proposed to model the evolution of software systems organized in
modules. It was proven formally to generate scale-free networks.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% Is there such a thing as softwareness? Can we reproduce it?

``softwareness''

Our hypothesis is that at least one of the presented models can synthesize
networks that resemble software networks. We already know that they produce
networks that, just like software networks, are scale-free. This single
property, though, isn't enough to prove the hypothesis, since there are many
known scale-free networks extracted from many distinct domains of knowledge. It
is important to distinguish between software and non-software networks.

% v.  differentiate, differ; distinguish; contrast; discern

The approach we chose to test our hypothesis was to build a network classifier:
an algorithm that analyses a network and classifies it as a software network or
as a non-software network. If a classifier is capable of classifying with high
accuracy both software and non-software networks...
it gives a fair approximation apply it on synthetic networks. We expect that
some synthetic networks will be so similar to software networks that they will
be classified as software networks.

The experiment we conducted to evaluate the models, described in details in the
following sections, can be summarized as such:
* First of all, we collected both software and non-software networks
* Then, we devised a classifier
* After that, we applied the classifier to our data set and evaluated its
accuracy, precision and recall.
* After we came to an acceptable classifier, we synthesized many networks by
varying the parameters on the models we presented.
* Finally, we classified all synthesized networks as software or non-software.

After that we analysed the results and tried to understand which parameters
of each model contribute mostly to the softwareness of their networks

%First of all, we need a data set containg both software and non-software
%networks. 
%Therefore, in order to prove our hypothesis we must show that, according to some
%criterion, some of at least some of
%the synthetic networks are more similar to software networks than other networks
%are.
%What we are looking for, after all, is an oracle that accepts software networks
%and rejects non-software networks. If the oracle has these two properties, we
%can be confident that it'll accept only synthetic networks that resemble
%software networks.

\subsection{Real Networks Data Set}

We've collected 65 systems written in Java and X networks from many domains,
such as sociology, biology, technology and linguistics.

See Appendix A

\subsection{Classifier} % Software network classifier

In a recent work, Milo et al. proposed to distinguish networks of different
domains by analyzing the triad concentration of each network. A triad is simply
a network with three vertices in which all vertices are connected. There are
only 13 distinct triads -- one for each configuration of directed edges --, as
shown in Figure X.
%CONSIDERING all the possible configurations of
%directed edges between three vertices, there are to 13 triads in which all
%vertices are connected, as shown in Figure X.

% Figure X: triad id vs. graphical representation

In order to characterize a network, one can count how many times each of the
13 triads appear on the network. The result is a triad concentration profile.
Figure 1a show triads for a software system... Figure 2a show triads for network
from domain X.

Then, in order to measure how similar two networks are, Milo et al computed
Pearson's correlation coefficient between the two corresponding profiles.

In order two measure the softwareness of a network, we measure the correlation
between the network's profile and the profile of each of the 65 software
networks in our sample. We then compute the average of these correlations and
call it the S-score, that varies between 0 and 1. The greater the S-score, more
likely that the network is a software network.

In order to build our classifier, we need a threshold for the S-score, so
networks are classified as software networks only when they have S-score greater
or equal to this threshold. We computed the S-score for each of the 65 software
systems and found an average S-score of 0.97 with standard deviation of 0.03.
Using the three-sigma rule from statistics, we chose the threshold as 0.97 - 3 *
0.3, that is 0.88.

The small standard deviation shows that triad concentration profiles indeed
characterize well the software domain.

\subsubsection{Classifier Validation}

In order to assess if the classifier is capable of distinguishing between
software and non-software networks, we applied it to all the networks from our
sample.

Most software networks were classified correctly, which is not surprising since
the classifier was based on them.

Precision: X, recall: Y

Curiously, polblog had a very high S-score. This issue requires further 
investigation. Anyway, we cannot deny the possibility that it is really very
software-like.

We could have chosen another value for the threshold, and it would affect the
precision and the recall. For instance, had we chosen 0.82, we would end with
100\% recall and x\% precision. For us, though, it is more important to have a
high precision, so we keep the threshold 0.88.

We also generated networks with the ER model, which is known to generate
networks that are not scale-free and, thus, are very different from software
networks. All the ER networks were classified as non-software.

%Following Milo et al., We then used Pearson's correlation coefficient as a
%similarity measure between two networks. For each software network, we then
%computed the average correlation coefficient (ACC) to the other software
%networks. We've observed that among software networks the ACC is X +- Y.

% avg: 0.97. stdev: 0.03
%We then computed, for each non-software network, the ACC to the 65 software
%networks. By the 3-sigma rule, we use X as the threshold for realistic software
%networks: networks whose ACC is below this threshold are rejected.
%The oracle has X precision and X recall...

@@

We chose five different module dependency networks, which where extracted from
actual dependencies between modules of five different software systems of our
sample, ranging from 2 to 32 modules. The module dependency networks are shown
on Figure XXX. 
%02-gef
%04-ibatis
%08-megamek
%16-findbugs
%32-zk

@@

Total: 9,500 networks.

\subsubsection{LF}

Like BCR, we choose mu ranging from 0.0 to 0.6 in 0.2 steps. For the remaining
parameters we selected values from our sample of software systems. For degexp,
..., we picked the minimum, the median and the max values.

Total: 1,296 networks

\subsubsection{CGW}

p1,p2,p3,p4 ranging from 0.0 to 1.0 in 0.2 steps, with p1 > 0, p1 + p2 + p3 + p4 = 1.0.
e1,e2,e3,e4 in 1, 2, 4, 8 (except that when pi = 0, ei is ignored).
2*p4*e4 >= p1*e1 + p2*e2, so the number of edges created is at least twice
the number of edges removed (to avoid long run times)
alpha in -1, 0, 1, 10, 100, 1000. 
m in 2, 4, 8, 16, 32 (just like bcr and lfr)

Total: 38,790 networks

\section{Softwareness Evaluation}

We used our classifier ...

Table: Model | Number of Networks | Realistic Networks | Percent \%

@@

@@

Pick realistic networks from a model. Are they similar to each other? (see
standard deviation) In other words: do the parameters make a difference?

Are they similar to networks generated by the other models? In other words: are
the models equivalent?

\section{Threats to validity}

@@

We generated only one network for each set of parameters. (as we've shown, some
parameters are redundant as they do not change significantly the realism)

@@

% Koschke says that experts decompositions vary by no more than 80? percent.

@@

We have shown empirically that network models found in the literature can
produce synthetic networks that are similar to the network of dependencies
between classes in object-oriented systems. This result supports the use of
synthetic networks in the evaluation of software modularization algorithms that
operate on class dependency networks.

% review
The use of synthetic data is common in distributed computing research, but still
underexplored in software engineering research. Since many reverse engineering
tasks extract from component dependencies information useful for the maintenance
of a software system, we expect this work to have impact beyond the software
modularization recovery community.

%Although the use of models and synthetic data is somewhat common in distributed
%computing research, it is underexplored in software engineering research. We
%expect this work to contribute to explore t
%common in distributed computing research, simulation is underexplored
%in software engineering. This work opens a door to the usage of simulation in
%the field of reverse engineering of software in order to evaluate RE 
%algorithms.

We accept that it is very important to test the algorithms with networks
extracted from real software systems, but we argue that the use of synthetic
networks in a complementary manner can give researchers new insights about the
algorithms. First, the use of models allows the creation of large test sets,
thus diminishing the small sample effects. Moreover, the networks are created in
a controlled way, accordingly to model parameters, so it is possible to study
the behavior of the algorithms with different parameter values.

In a future work, we intend to evaluate with synthetic networks some software
clustering algorithms that were previously studied with real data. We will,
then, be able to compare the results obtained with the two approaches.

@@

From SourceForge:
AbaGuiBuilder-1.8
alfresco-labs-deployment-3Stable
aoi272
stendhal-0.74
battlefieldjava-0.1
checkstyle-5.0
dom4j-1.6.1
findbugs-1.3.8
freetts-1.2.2-bin
ganttproject-2.0.9
geoserver-2.0-beta1-bin
geotools-2.5.5-bin
gfp\_0.8.1
hibernate-distribution-3.3.1.GA-dist
hsqldb\_1\_8\_0\_10
iBATIS\_DBL-2.1.5.582
iReport-nb-3.5.1
JabRef-2.5b2-src
jailer\_2.9.9
jalopy-1.5rc3
jasperreports-3.5.2-project
jfreechart-1.0.13
pentaho-reporting-engine-classic-0.8.9.11
jGnash-2.2.0
jgraphpad-5.10.0.2
jmsn-0.9.9b2
juel-2.1.2
JXv3.2rc2deploy
makagiga-3.4
MegaMek-v0.34.3
iFreeBudget-2.0.9
mondrian-3.1.1.12687
oddjob-0.26.0
openxava-3.1.2
pdfsam-1.1.3-out
pjirc\_2\_2\_1\_bin
pmd-bin-4.2.5
proguard4.3
smc\_6\_0\_0
squirrel-sql-3.0.1-base
squirrel-sql-3.0.1-standard
tvbrowser-2.7.3-bin
villonanny-2.3.0.b02.bin
rapidminer-4.4-community
zk-bin-3.6.1

From other places:

ArgoUML-0.28
GEF-0.13-bin
Hl7Comm.1.0.1
IRPF2009v1.1
broker-4.1.5
dbwrench
ec2-api-tools
ermodeller-1.9.2-binary
flyingsaucer-R8
gdata-src.java-1.31.1
guice-2.0
gwt-windows-1.6.4
jai-1\_1\_4-pre-dr-b03-lib-linux-i586-08\_Jun\_2009
jakarta-tomcat-5.0.28-embed
juxy-0.8
myjgui\_0.6.6
peer-4.1.5
subethasmtp-3.1
thinkui\_sqlclient-1.1.2
worker-4.1.5

@@

3 circuit networks (circuit-s208 circuit-s420 circuit-s838)

5 facebook networks (facebook-Caltech36 facebook-Georgetown facebook-Oklahoma
facebook-Princeton facebook-UNC28)

5 language networks (lang-english  lang-french lang-japanese lang-spanish)

43 metabolic networks

polblogs

3 protein networks (protein-a4j protein-AOR protein-eaw)

2 social networks (social-leader social-prison)

other networks:

yeast

beta3sreduced
celegansneural

czech
ecoli-metabolic
